---
title: "Final"
author: "Adam Guerra, Luke Todd, Riley Zamora"
date: "2023-06-11"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 5
    theme: sandstone
    code_folding: TRUE
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading Packages
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) 
library(flextable) 
library(car)
library(broom)
library(corrplot)
library(AICcmodavg)
library(MASS)
library(lme4)
library(MuMIn)
library(lmtest)
library(DHARMa)
library(ggeffects)
library(lmtest)
library(skimr)
library(GGally)
library(glmmTMB)
library(broom.mixed)
```


## Problem 1


### Introduction


### Methods

##### Loading Data
```{r, warning = FALSE}
# loading data
kang_seed <- read.table(here("data/sev208_kratseedbank_20120213.txt"), sep=',', header = TRUE) |> 
             mutate(mnd = as.factor(mnd)) |> 
             filter(!(species %in% c("soil", "plant", "dist", "litter", "gravel")))
```

##### Missing Data
```{r}
gg_miss_var(kang_seed)
```

##### Skim
```{r}
skim(kang_seed)
```

##### Visualize Data
```{r}
# visualize counts
ggplot(kang_seed, aes(x = seeds)) +
  geom_histogram(bins = 17)

```

```{r}
#visualize using box plot
ggplot(data = kang_seed, aes(x = mnd, y = seeds)) +
  geom_boxplot() +
  labs(x = "Mound Number", y = "NUmber of Seeds", title = "Boxplot of Seeds Found at Different Mounds") +
  theme_bw()
```

```{r}
# set up data for bar graph
kang_seed_bar <- kang_seed |> 
                 group_by(mnd) |> 
                 summarise(across(seeds, sum)) |> 
                 ungroup()
# bar graph
ggplot() +
  geom_bar(data = kang_seed_bar, aes(x = mnd, y = seeds, fill = mnd), 
           stat = "identity", show.legend = F) +
  theme_bw() +
  labs(x = "Mound Number", y = "Number of Seeds", title = "Total Seed Count per Mound") 
```

**Question: How does total seed number differ between kangaroo rat mound locations?**

##### Choose Test  
We will be performing an ANOVA test to determine if there is a difference in the total seed count between kangaroo rat mound locations.

##### Set Up Hypothesis  
$H_0:$ There is no significant difference in seed counts between the kangaroo rat mound locations.
$H_a:$ There is a significant difference in seed counts between at least two of the kangaroo mound locations.

##### Initial Assumption Checks
```{r}
# make aov object
kang_aov <- aov(seeds ~ mnd, data = kang_seed)

# get resiudals
kang_res <- kang_aov$residuals

# check normality
qqPlot(kang_res)

```

The data does not appear to be normal, which we expceted because in the context reading they discussed how in their analysis the data was non-normal even after trying transformations. We decided to try a log transformation regaudless because this would be the logical next step in this analysis.

```{r}
# transform data
kang_seed_log <- kang_seed |> 
                 mutate(seeds = log(seeds + 1))

# make new aov object
kang_log_aov <- aov(seeds ~ mnd, data = kang_seed_log)

# residuals
kang_log_res <- kang_log_aov$residuals

#test normality
qqPlot(kang_log_res)
```

The data is still non-normal as expected (we mentionded above that the context paper mentioned how the data is not normal), we will now test the variance of the un-transformed data.

```{r}
# check variance using levenes test
leveneTest(seeds ~ mnd, data = kang_seed)
```

After running the levenes test, we fail to reject the null hypothesis that the variances between groups are equal. Below we show the reuslts from the ANOVA test, we are still able to run it can handle non-normal data and we met the other assumptions.

```{r}
# show aov summary
summary(kang_aov)
```

The ANOVA p-value is greater than 0.05, meaning we fail to reject the null hypothesis that there is no significant difference in total seed count between different kangaroo rat mounds.

### Results



---

## Problem 2


### Introduction



### Methods


##### Loading Data
```{r}
# loading data
shrub_raw <- read.csv(here("data/shrubstudy_seed.csv"))

shrub_seed <- shrub_raw |>
              dplyr::select(treatment, species, total_inf = total_nr_infl, 
                      num_seeds = nr_seeds, shrub_num, plant_num = plant_nr, tag_num) |>
              mutate(treatment = as.factor(treatment))
```

##### Missing Data
```{r}
# visualize missing data
gg_miss_var(shrub_seed)
```

##### Skim
```{r}
skim(shrub_seed)
```

##### Subset Dataset
```{r}
# filter out missing data
shrub_seed_sub <- shrub_seed |> 
                  drop_na(num_seeds) 
```

##### Histogram of Counts
```{r}
ggplot(shrub_seed_sub, aes(x = num_seeds)) +
  geom_histogram(bins = 17)
```

##### Correlation Plot
```{r}
# calc Pearson's r for numerical values
shrub_seed_sub_cor <- shrub_seed_sub |> 
                      dplyr::select(3:4) |> 
                      cor(method = "pearson")

# plot correlation values
corrplot(shrub_seed_sub_cor,
         method = "ellipse",
         addCoef.col = "black")
```
    
##### Variable Relationships
```{r, warning=FALSE, message=FALSE}
shrub_seed_sub |> 
  dplyr::select(!num_seeds) |> 
  ggpairs()
```

**Question: How does seed count vary with plot type (shrub or open), plant species, and total number of inflorescences? Is there a simpler model that explains seed count, and if so, what is it?**

##### Build Models
```{r, message=FALSE, warning=FALSE}
# linear model, we know this is wrong
shrub_model1 <- lm(num_seeds ~ treatment + species + total_inf, data = shrub_seed_sub)

# generalized linear model with Poisson distribution
shrub_model2 <- glm(num_seeds ~ treatment + species + total_inf, data = shrub_seed_sub, family = "poisson")

# generalized linear model with negative binomial distribution
shrub_model3 <- glmmTMB(num_seeds ~ treatment + species + total_inf, data = shrub_seed_sub, family = "nbinom2")

# generalized linear model with Poisson distribution and random effect of treatment
shrub_model4 <- glmmTMB(num_seeds ~ treatment + species + total_inf + (1|shrub_num), 
                          data = shrub_seed_sub, family = "poisson")

# generalized linear model with negative binomial distribution and random effect of treatment
shrub_model5 <- glmmTMB(num_seeds ~ treatment + species + total_inf + (1|shrub_num), 
                      data = shrub_seed_sub, family = "nbinom2")
```

Because we are looking at count data, we know that the data is discrete and only has a lower bound. Knowing this, we built a couple different models using the Poisson and Negative Binomial distribution.

```{r}
# check diagnostics
simulationOutput_m1 <- simulateResiduals(shrub_model1)

plot(simulationOutput_m1) # bad
testDispersion(simulationOutput_m1)
```
```{r}
# check diagnostics
simulationOutput_m2 <- simulateResiduals(shrub_model2)

plot(simulationOutput_m2) # bad
testDispersion(simulationOutput_m2)
```
```{r}
simulationOutput_m3 <- simulateResiduals(shrub_model3)

plot(simulationOutput_m3) # bad
testDispersion(simulationOutput_m3)
```
```{r}
# check diagnostics
simulationOutput_m4 <- simulateResiduals(shrub_model4)

plot(simulationOutput_m4) # bad
testDispersion(simulationOutput_m4)
```
```{r}
# check diagnostics
simulationOutput_m5 <- simulateResiduals(shrub_model5)

plot(simulationOutput_m5) # bad
testDispersion(simulationOutput_m5)
```

##### Which distribution to use?
```{r}
MuMIn::model.sel(shrub_model1, shrub_model2, shrub_model3, shrub_model4, shrub_model5)
```

##### Model Summary
```{r}
# model object
shrub_model3
```
```{r}
# summary
summary(shrub_model3)
```

```{r, warning=FALSE, message = FALSE}
# confidence intervals
confint(shrub_model3)
```
```{r, warning = FALSE}
# adjusted R2
r.squaredGLMM(shrub_model3)
```
```{r}
# model object in table
shrub_model3 |> 
  as_flextable()
```

##### Visualizing
```{r}

```



### Results
































